---
title: "net for rfp"
author: "Clint McKenna"
date: "February 20th, 2020"
output: html_document
---

## data prep
```{r}
# load packages
library(igraph)
library(tidyverse)
library(ggraph)


# custom functions
# dplyr select
select <- dplyr::select

# read in data seperately
dat <- read_csv('../data/data.csv')
rounds <- read_csv('../data/rounds.csv')
edgelist <- read_csv('../data/edgelist.csv')


```


## prepare network object
```{r}

edgelist <- data.frame(
  sender = c(
    rep('A', 4),
    rep('B', 4),
    rep('C', 4),
    rep('D', 4),
    rep('E', 4),
    rep('F', 4),
    rep('G', 4),
    rep('H', 4),
    rep('I', 4),
    rep('J', 4),
    rep('K', 4),
    rep('L', 4)),
  receiver = c(
    'K', 'L', 'B', 'C',
    'L', 'A', 'C', 'D',
    'A', 'B', 'D', 'E',
    'B', 'C', 'E', 'F',
    'C', 'D', 'F', 'G',
    'D', 'E', 'G', 'H',
    'E', 'F', 'H', 'I',
    'F', 'G', 'I', 'J',
    'G', 'H', 'J', 'K',
    'H', 'I', 'K', 'L',
    'I', 'J', 'L', 'A',
    'J', 'K', 'A', 'B')
)


# for the next steps, we need to move title to the last column
edge1 <- edgelist %>%
  select(sender, receiver)

# retrieve vertex scores from initial dataset
vertexScores <- data.frame(
  role = LETTERS[1:12],
  score = c(rep(0, 3), 1, rep(0, 8)))

  
# create color palette function for each score
# vertexScores$vColors <- c('red', 'white', 'red', 'white', 'red', 'white')
colR <- colorRampPalette(c('red', 'white'))
# assign it to an object with 6 colors
colPal <- colR(2)

# assign color gradient depending on r1_score
vertexScores <- vertexScores %>%
  mutate(
    colors = case_when(
      role == 'D' ~ colPal[2],
      TRUE ~ colPal[1]))



# sort edgelist by sender
edge1 <- edge1[order(edge1$sender),]
vertexScores <- vertexScores[order(vertexScores$role),]


# create network object. non-directed
net <- graph_from_data_frame(
  edge1,
  directed = FALSE,
  vertices = vertexScores
)



```


## preview one time point
```{r}
plot(net,
     layout = layout_in_circle,
     # layout = layout_nicely,
     vertex.label.cex = 2,
     # vertex.label.family = 'Times',
     # vertex.label.font=20,
     vertex.label.color='black',
     # vertex.label = NA,
     # vertex.shape='circle', 
     # vertex.size = degree(net)*2,
     vertex.size = 20,
     # vertex.color = 'light blue',
     # vertex.color = vertex_attr(net)$r1_score,
     vertex.color = vertex_attr(net)$r1_colors,
     edge.curved = FALSE,
     edge.width = 2)

```



## plot all time points
make sure to run this entire code chunk at once
```{r}

# set file name, dimensions, and resolution
png('../figures/net.png', units = 'in', width = 16, height = 4, res = 750)

# arrage 4 plots in a line
par(mfrow=c(1,4), family = 'Sen')

# baseline (time 1)
plot(net,
     layout = layout_in_circle,
     vertex.label.cex = 2,
     vertex.label.color='black',
     vertex.size = 75,
     vertex.color = vertex_attr(net)$r1_colors,
     edge.curved = FALSE,
     edge.width = 2)
title(main = 'Baseline', cex.main = 2.5)

# round 1 (time 2)
plot(net,
     layout = layout_in_circle,
     vertex.label.cex = 2,
     vertex.label.color='black',
     vertex.size = 75,
     vertex.color = vertex_attr(net)$r2_colors,
     edge.curved = FALSE,
     edge.width = 2)
title(main = 'Round 1', cex.main = 2.5)

# Round 2 (time 3)
plot(net,
     layout = layout_in_circle,
     vertex.label.cex = 2,
     vertex.label.color='black',
     vertex.size = 75,
     vertex.color = vertex_attr(net)$r3_colors,
     edge.curved = FALSE,
     edge.width = 2)
title(main = 'Round 2', cex.main = 2.5)

# round 3 (time 4)
plot(net,
     layout = layout_in_circle,
     vertex.label.cex = 2,
     vertex.label.color='black',
     vertex.size = 75,
     vertex.color = vertex_attr(net)$r4_colors,
     edge.curved = FALSE,
     edge.width = 2)
title(main = 'Round 3', cex.main = 2.5)

# save to file
dev.off()

```






## ggraph package
```{r}


p <- ggraph(net, layout = 'circle') +
  geom_edge_link(alpha = .25) +
  geom_node_point(aes(color = V(net)$score), size = 10, alpha = .95) +
  geom_node_text(aes(label = names(V(net))), color = 'white') +
  scale_color_gradient(low = '#3b5998', high = 'red') +
  labs(
    title = 'Network Structure',
    color = 'Response Accuracy') +
  theme_graph() +
  theme(legend.position = 'none') + 
  coord_cartesian(xlim=c(-1.2,1.2), ylim=c(-1.2,1.2))






png('../figures/net.png', units = 'in', width = 6, height = 6, res = 500)
p
dev.off()




```





The nice thing about ggplot2 based plots is that its easier to facet. Lets make a net object with all the data and facet in ggraph.
```{r}


# for the next steps, we need to move title to the last column
edge1 <- edgelist %>%
  select(sender, receiver, condition, subject)

# rename participant roles for each condition, for both sender and receiver
edge1 <- edge1 %>%
  mutate(
    sender = case_when(
      condition == 'visible' ~ gsub('Participant ', 'visible_', .$sender),
      condition == 'notVisible' ~ gsub('Participant ', 'notVisible_', .$sender)),
    receiver = case_when(
      condition == 'visible' ~ gsub('Participant ', 'visible_', .$receiver),
      condition == 'notVisible' ~ gsub('Participant ', 'notVisible_', .$receiver)))

# retrieve vertex scores from initial dataset
vertexScores <- dat %>%
  select(sender = role, condition, r1_score, r2_score, r3_score, r4_score)
# rename participant roles for each condition
vertexScores <- vertexScores %>%
  mutate(sender = case_when(
    condition == 'visible' ~ gsub('Participant ', 'visible_', .$sender),
    condition == 'notVisible' ~ gsub('Participant ', 'notVisible_', .$sender)))



# sort edgelist by sender
edge1 <- edge1[order(edge1$sender),]
vertexScores <- vertexScores[order(vertexScores$sender),]

# only edgelist vals
edge2 <- edge1 %>% select(sender, receiver)

# create network object. non-directed
net <- graph_from_data_frame(
  edge2,
  directed = FALSE,
  vertices = vertexScores
)


```



```{r}

ggraph(net, layout = 'circle') +
  geom_edge_link() +
  geom_node_point(aes(color = V(net)$r1_score), size = 10, alpha = .75) +
  scale_color_gradient(low = 'red', high = 'light blue') +
  labs(
    title = 'Baseline',
    color = 'Response Accuracy') +
  theme_graph() +
  facet_grid(~ V(net)$condition)

```

No idea why it is preserving the nodes not present in one condition over another...



## patchwork
Another option is using patchwork package. Here, we will "glue" a bunch of plots together.
```{r}

# visible condition
visible_edge1 <- edge1 %>% 
  filter(condition == 'visible') %>%
  select(sender, receiver)
visible_vertexScores <- vertexScores %>%
  filter(condition == 'visible')
# create network object. non-directed
netVisible <- graph_from_data_frame(
  visible_edge1,
  directed = FALSE,
  vertices = visible_vertexScores
)



# notVisible condition
notVisible_edge1 <- edge1 %>% 
  filter(condition == 'notVisible') %>%
  select(sender, receiver)
notVisible_vertexScores <- vertexScores %>%
  filter(condition == 'notVisible')
# create network object. non-directed
netNotVisible <- graph_from_data_frame(
  notVisible_edge1,
  directed = FALSE,
  vertices = notVisible_vertexScores
)

```

## plot
```{r}

library(patchwork)

# visible
p1 <- ggraph(netVisible, layout = 'circle') +
  geom_edge_link() +
  geom_node_point(aes(color = V(netVisible)$r1_score), size = 10, alpha = .75) +
  scale_color_gradient(low = 'red', high = 'light blue') +
  labs(
    title = 'Visible',
    color = 'Response Accuracy') +
  theme_graph() +
  theme(legend.position = 'none')

# not visible
p2 <- ggraph(netNotVisible, layout = 'circle') +
  geom_edge_link() +
  geom_node_point(aes(color = V(netNotVisible)$r1_score), size = 10, alpha = .75) +
  scale_color_gradient(low = 'red', high = 'light blue') +
  labs(
    title = 'Not Visible',
    color = 'Response Accuracy',
    caption = 'Grey nodes indicate NA values') +
  theme_graph()


# output both plots
p1 + p2


# can also save to file in same way as before (run these lines altogether)
png('../figures/ggnet.png', units = 'in', width = 12, height = 4, res = 750)
p1 + p2
dev.off()

```






